---
title: Analyzing US Census Data
author: David Kane, Ansh Pathak, and Paramanyu Gupta
tutorial:
  id: census
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Companion to the tidycensus chapters of 'Analyzing US Census Data' by Kyle Walker. Pulls American Community Survey data with tidycensus and builds three published artifacts: a choropleth map of Texas household income, a ranked chart of state educational attainment, and a scatterplot of California county population and median age."
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(tidycensus)
library(sf)
library(knitr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

# The data below was pulled once with tidycensus and saved, so this tutorial
# renders without hitting the Census API. Students make the live calls in their
# own analysis.qmd.

data_path <- function(file) {
  input <- knitr::current_input(dir = TRUE)
  candidates <- c(
    file,
    if (is.character(input)) file.path(dirname(input), file),
    file.path("inst/tutorials/census", file)
  )

  for (candidate in candidates) {
    if (file.exists(candidate)) {
      return(candidate)
    }
  }

  file
}

# income_tx <- get_acs(geography = "county", variables = "B19013_001",
# state = "TX", year = 2020, geometry = TRUE)
# write_rds(income_tx, "data/income_tx.rds")
income_tx <- read_rds(data_path("data/income_tx.rds"))

# edu_state <- get_acs(geography = "state",
# variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
# year = 2020, geometry = FALSE, summary_var = "B15003_001")
# write_rds(edu_state, "data/edu_state.rds")
edu_state <- read_rds(data_path("data/edu_state.rds"))

# age_ca <- get_acs(geography = "county",
# variables = c(median_age = "B01002_001", population = "B01003_001"),
# state = "CA", year = 2020, geometry = FALSE)
# write_rds(age_ca, "data/age_ca.rds")
age_ca <- read_rds(data_path("data/age_ca.rds"))
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial follows Kyle Walker's [An introduction to tidycensus](https://walker-data.com/census-r/an-introduction-to-tidycensus.html) and [Mapping Census data with R](https://walker-data.com/census-r/mapping-census-data-with-r.html) chapters from [Analyzing US Census Data](https://walker-data.com/census-r/). You will pull data from the American Community Survey through the Census API and turn it into three published artifacts: a choropleth map of Texas household income, a ranked chart of state educational attainment, and a scatterplot of California county population and median age.

Use AI however you like throughout. An agentic coding tool (such as Claude Code or Gemini CLI) works best, since it can read and edit your files directly; a chat-based assistant works too, with more copy/paste.

### Exercise 1

You should be connected to a repo named `census`. If you are not, create one and connect to it. Create a new file and save it as `analysis.qmd`. Add a YAML header with a title (`"Analyzing US Census Data"`) and your name as author. In a bash Terminal, run:

```
quarto render analysis.qmd
```

Open `analysis.html` by right-clicking it in the Explorer and selecting **Open with Live Server**. Keep this tab open — it refreshes automatically whenever you re-render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
analysis_files
```

###

The American Community Survey (ACS) estimates characteristics like income, education, and age from a sample. That is why ACS values come with margins of error.

### Exercise 2

In `analysis.qmd`, add a code chunk that loads the three libraries you will need:

```
library(tidyverse)
library(tidycensus)
library(sf)
```

Rendering shows the code and the package start-up messages, which we don't want in a published document. Add `#| message: false` to the chunk, and add this to the YAML header to hide all code:

```
execute:
  echo: false
```

Render again. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
library(tidyverse)
library(tidycensus)
library(sf)</code></pre>

###

**tidycensus** returns tidy ACS tables. When you request geometry, it returns an **[sf](https://r-spatial.github.io/sf/)** object that **[ggplot2](https://ggplot2.tidyverse.org/)** can map directly.

### Exercise 3

Every call to `get_acs()` needs a Census API key. Sign up for a free one at the [Census API key signup page](https://api.census.gov/data/key_signup.html) — the key arrives by email within a few minutes.

In the R Terminal, run (with your own key in place of `"YOUR-KEY-HERE"`):

```
tidycensus::census_api_key("YOUR-KEY-HERE", install = TRUE)
```

<!-- AR: Look at the .Renviron file in exercise 4. CP/CR but delete the key. Never share your API keys publicly. KD: WARNING, doesn't set key for all CODESPACEs-->

This stores the key in your `.Renviron` file so every future R session finds it automatically. In the bash Terminal, run:

```
cat ~/.Renviron
```

You should see a line that starts with `CENSUS_API_KEY=`. Do not paste that output anywhere, because it contains your key.

<!-- AR: Check if resetting R session is needed. Test in codespaces. Below should end exercise 3. -->

Restart your R session.

```
nzchar(Sys.getenv("CENSUS_API_KEY"))
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
[1] TRUE
```

###

The key lives in `.Renviron`, not in `analysis.qmd`. That keeps it out of your code and commits.

## Texas income
###

Median household income is one of the most-requested ACS variables. Its code is `B19013_001`. In this section you will map it across all 254 Texas counties.

### Exercise 1

<!-- AR: Split into 3 exercises. 1. Run get acs. Vomits out data frame. 2. save to a variable. 3. Add message: false -->

Add a chunk to `analysis.qmd` that uses **tidycensus** to get 2020 median household income for every county in Texas, with county geometry attached for mapping, and saves it as `income_tx`. Add `#| message: false` to the chunk to hide the connection messages. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
income_tx <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "TX",
  year = 2020,
  geometry = TRUE
)
</code></pre>

###

<!-- AR: Prompt Codex to make all KDs a max of 2 sentences. -->

Every ACS request needs a geography, variables, and a year. For ACS data, `year = 2020` means the pooled 2016-2020 5-year estimates.

### Exercise 2

Add a chunk that prints `income_tx` so you can see its structure. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r texas-income-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r texas-income-2-test}
#| echo: true
income_tx
```

###

Every row has an `estimate` and a margin of error (`moe`). Because you asked for boundaries, it also has a geometry column for the county polygon.

### Exercise 3

Replace the working chunk with code that shows the structure of `income_tx`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r texas-income-3-test}
#| echo: true
glimpse(income_tx)
```

###

The `geometry` column is the giveaway that `income_tx` is an **sf** object. It is still a table, but each row carries a county polygon.

### Exercise 4

Replace the working chunk with code that summarizes median household income and margin of error across the Texas counties. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

```{r texas-income-4-test}
#| echo: true
income_tx |>
  st_drop_geometry() |>
  summarize(
    min_income = min(estimate),
    median_income = median(estimate),
    max_income = max(estimate),
    median_moe = median(moe)
  )
```

###

The income spread is wide enough that a map should be interesting. Drop geometry when the calculation is purely tabular.

### Exercise 5

Replace the working chunk with a histogram of county median household income. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r texas-income-5-test}
#| echo: true
#| message: false
ggplot(income_tx, aes(x = estimate)) +
  geom_histogram(bins = 30, fill = "#2b8cbe", color = "white") +
  labs(
    title = "Distribution of median household income across Texas counties",
    subtitle = "Each bar counts counties by their ACS median household income estimate",
    x = "Median household income",
    y = "Number of counties"
  ) +
  theme_minimal()
```

###

The histogram shows the spread, but not where the counties are. The map will show whether high-income counties cluster geographically.

### Exercise 6

<!-- AR: Does this instruction work as a prompt? -->

Replace the histogram with a choropleth map of median income across the counties using **ggplot2**. Render and check the map in your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<!-- AR: Is this a chloropleth? Needs axis labels, titles, subtitles etc... -->

```{r texas-income-6-test}
#| echo: true
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate)) +
  labs(
    title = "Median household income by county, Texas (2020)",
    subtitle = "A first choropleth reveals where high- and low-income counties cluster",
    x = "Longitude",
    y = "Latitude",
    fill = "Median income"
  )
```

###

This is a **choropleth**: each county is shaded by a value. The first version works, but the default colors and gridlines add clutter.

### Exercise 7

Improve the map: give it a perceptually uniform color scale (the **[viridis](https://sjmgarnier.github.io/viridis/)** family is the standard choice), a clear title and legend label, a caption naming the data source, and a clean map theme. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r texas-income-7-test}
#| echo: true
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate), color = "white", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Median income") +
  labs(
    title = "Median household income by county, Texas (2020)",
    subtitle = "Suburban counties around major metros stand out from much of rural Texas",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
```

###

The finished map makes the metro-suburban pattern visible at a glance. A choropleth is strongest when the geography carries the message.

### Exercise 8

Commit your work. Stage `analysis.qmd`, commit with the message `"Add Texas income map"`, and push to GitHub. You can do this with your AI agent, the VS Code Source Control panel, or `git` on the command line.

CP/CR.

```{r texas-income-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Median household income is a good choropleth variable because it is comparable across counties. Raw counts often just map population.

## Education by state
###

Educational attainment lives in ACS table B15003, which breaks the adult population into 25 completed schooling levels from "no schooling" up to "doctorate." In this section you will compare the share of adults with at least a bachelor's degree across all state-level geographies.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 educational attainment for state-level geographies and save it as `edu_state`. You need the table total plus the bachelor's-and-higher levels: variables `B15003_001` (total), `B15003_022`, `B15003_023`, `B15003_024`, and `B15003_025`. Pass `summary_var = "B15003_001"` so the total rides along on every row. Add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

<pre><code>#| message: false
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
</code></pre>

###

These codes report completed degrees, not enrollment. Adding bachelor's, master's, professional, and doctorate counts gives the numerator for "at least a bachelor's."

### Exercise 2

Add a working chunk that prints `edu_state`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r education-by-state-2-test}
#| echo: true
edu_state
```

###

The result is **long**: each geography appears once per requested variable. Here that means 52 geographies times five variables, or 260 rows.

<!-- AR: In a new code chunk run edu-state to look at it. -->

### Exercise 3

Replace the working chunk and run `glimpse(edu_state)`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r education-by-state-3-test}
#| echo: true
glimpse(edu_state)
```

###

The important columns are `estimate`, the count for one education level, and `summary_est`, the repeated adult total. The margin-of-error columns remind you these are ACS estimates.

### Exercise 4

Replace the working chunk with a summary by `variable` that reports how many rows each variable has and the smallest and largest estimates. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-4-test}
#| echo: true
edu_state |>
  summarize(
    rows = n(),
    min_estimate = min(estimate),
    max_estimate = max(estimate),
    .by = variable
  )
```

###

Each variable has 52 rows, one per state-level geography. Raw counts would mostly reward large populations, so the final comparison needs percentages.

### Exercise 5

You are done changing the chunk that downloads `edu_state`, and re-downloading it on every render is slow. `edu_state` earns its own cached create chunk. The next three exercises set that up: clean the chunk, turn caching on, add the cache directory to `.gitignore`.

First, clean up. Delete the working chunk that summarized `edu_state`, leaving the `edu_state <- get_acs(...)` chunk as the last chunk in the file. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r education-by-state-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

<pre><code>#| message: false
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
</code></pre>

###

A cached chunk should contain only the artifact it builds. The exploratory chunk helped you understand the data, but it does not need to stay.

### Exercise 6

Add `#| cache: true` as the second line inside the chunk, just below `#| message: false`. Render the QMD. Rendering creates an `analysis_cache/` directory next to your `analysis.qmd`, which holds the saved tibble so Quarto can reload it on subsequent renders rather than re-downloading from the Census API. To confirm the directory is there, run `ls` in the **bash Terminal**. CP/CR.

```{r education-by-state-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis.qmd  analysis_cache  analysis_files  analysis.html
</code></pre>

###

Caching belongs on finished, expensive chunks. The cache is useful locally, but it is generated output.

### Exercise 7

The Source Control change count just jumped, because `analysis_cache/` and everything inside it now look like untracked work to git. Cached files should never go to GitHub.

Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r education-by-state-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis_files
analysis_cache
</code></pre>

###

The Source Control count should drop after `.gitignore` catches the cache. Generated output can help locally without becoming source code.

### Exercise 8

Add a new working chunk that collapses `edu_state` to one row per geography. Compute the percentage of adults with at least a bachelor's degree, then print the table sorted from highest to lowest percentage. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-8-test}
#| echo: true
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    adults = unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  arrange(desc(percent))
```

###

The District of Columbia sits far above the states because it is a city-like jurisdiction. Always read the geography before comparing places.

### Exercise 9

Replace the working chunk with code that summarizes the distribution of the state-level bachelor's-or-higher percentages. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-9-test}
#| echo: true
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  pull(percent) |>
  summary()
```

###

The middle geography is around 31%, with a much higher maximum. A plot makes that shape easier to judge than a summary.

### Exercise 10

Replace the working chunk with a histogram of the state-level bachelor's-or-higher percentages. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-10-test}
#| echo: true
#| message: false
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  ggplot(aes(x = percent)) +
  geom_histogram(bins = 15, fill = "#756bb1", color = "white") +
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  labs(
    title = "Distribution of bachelor's-or-higher rates across state-level geographies",
    subtitle = "Most state-level geographies cluster near the middle, with DC far to the right",
    x = "Share of adults 25 and over",
    y = "Number of geographies"
  ) +
  theme_minimal()
```

###

Most geographies cluster in the middle, with DC far to the right. A ranked chart will identify every place by name.

### Exercise 11

Replace the working chunk with a horizontal bar chart of the bachelor's-or-higher percentage for every state-level geography, sorted by percentage. Render and look at it.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 9)
```

###

```{r education-by-state-11-test}
#| echo: true
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Adults with a bachelor's degree or higher by state-level geography (2020)",
    subtitle = "The rough ranking shows every geography before final formatting",
    x = "State-level geography",
    y = "Share of adults 25 and over"
  )
```

###

Ordering the bars turns the distribution into a comparison. The rough chart has the ranking; the final chart needs polish.

### Exercise 12

Improve the chart: add a clear title, axis label, source caption, percent-formatted axis labels, and a cleaner theme. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 9)
```

###

```{r education-by-state-12-test}
#| echo: true
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent, fill = percent)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Adults with a bachelor's degree or higher by state-level geography (2020)",
    subtitle = "DC and several Northeastern states sit above most of the country",
    x = "State-level geography",
    y = "Share of adults 25 and over",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
```

###

The final chart is a percentage comparison, not a population-size ranking. DC and several Northeastern states sit well above most places.

### Exercise 13

Commit your work. Stage `analysis.qmd`, commit with the message `"Add state education chart"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r education-by-state-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

The variables define what you measure. The geography defines the units you compare.

## California median age
###

Not every Census question is a map. Here you will pull median age and total population for California counties, then ask whether larger counties tend to be younger or older. This section also revisits the reshaping step that long-format ACS data often needs.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 median age and total population for every county in California, naming the variables `median_age` (`B01002_001`) and `population` (`B01003_001`). You are building a non-map chart, so set `geometry = FALSE`. Save it as `age_ca` and add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
age_ca <- get_acs(
  geography = "county",
  variables = c(median_age = "B01002_001", population = "B01003_001"),
  state = "CA",
  year = 2020,
  geometry = FALSE
)
</code></pre>

###

Readable variable names make the rest of the code easier to audit. Since this section does not map counties, `geometry = FALSE` avoids downloading polygons.

### Exercise 2

Add a working chunk that prints `age_ca`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r california-median-age-2-test}
#| echo: true
age_ca
```

###

Again the data is **long**: each county has one row per variable. A scatterplot needs those variables side by side.

### Exercise 3

<!-- AR: explicit glimpse -->

Replace the working chunk with code that shows the structure of `age_ca`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r california-median-age-3-test}
#| echo: true
glimpse(age_ca)
```

###

The `variable` column tells you which measure a row holds. The `estimate` column holds the value.

### Exercise 4

Replace the working chunk with a summary by `variable` that reports how many rows each variable has and the smallest and largest estimates. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-4-test}
#| echo: true
age_ca |>
  summarize(
    rows = n(),
    min_estimate = min(estimate),
    max_estimate = max(estimate),
    .by = variable
  )
```

###

Each variable has 58 rows, one per California county. Population varies much more than median age.

### Exercise 5

<!-- AR: Flow is muddled here. Run through the tutorial by hand and fix. -->

Replace the working chunk with code that reshapes `age_ca` to one row per county, with one column for `median_age` and one for `population`. Print the table sorted from largest to smallest population. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

```{r california-median-age-5-test}
#| echo: true
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  arrange(desc(population))
```

###

**Reshaping** turns repeated variable/value rows into one column per variable. That gives you one row per county.

### Exercise 6

Replace the working chunk with a histogram of median age across California counties. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-6-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  ggplot(aes(x = median_age)) +
  geom_histogram(bins = 15, fill = "#41ab5d", color = "white") +
  labs(
    title = "Distribution of median age across California counties",
    subtitle = "Most counties are in the mid-30s to mid-40s, with a few older outliers",
    x = "Median age",
    y = "Number of counties"
  ) +
  theme_minimal()
```

###

Most counties sit in the mid-30s to mid-40s. The next question is whether older counties are large or small.

### Exercise 7

Replace the working chunk with a scatterplot of county population against median age. Render and look at the relationship.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-7-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point() +
  labs(
    title = "County population and median age, California (2020)",
    subtitle = "A rough scatterplot shows the relationship before scale and label refinements",
    x = "Population",
    y = "Median age"
  )
```

###

The rough scatterplot has the right variables, but population is too skewed for a linear scale. A log scale gives smaller counties room to breathe.

### Exercise 8

Improve the scatterplot: use a log-scaled population axis, label the largest counties, add a title, clean axis labels, and a source caption. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r california-median-age-8-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  mutate(label = if_else(min_rank(desc(population)) <= 8, NAME, "")) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point(color = "#2c7fb8", alpha = 0.75, size = 2.5) +
  geom_text(aes(label = label), hjust = -0.05, check_overlap = TRUE, size = 3) +
  scale_x_log10(
    labels = scales::label_number(scale_cut = scales::cut_short_scale()),
    expand = expansion(mult = c(0.02, 0.25))
  ) +
  labs(
    title = "Smaller California counties tend to be older (2020)",
    subtitle = "The largest counties are mostly younger than California's smallest counties",
    x = "Population (log scale)",
    y = "Median age",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
```

###

The final plot shows a negative relationship: many of the oldest counties are small. The log scale keeps Los Angeles visible without crushing the smaller counties.

### Exercise 9

Commit your work. Stage `analysis.qmd`, commit with the message `"Add California age scatterplot"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r california-median-age-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Median age is an ACS estimate too. For small counties, a one- or two-year difference may be inside the margin of error.

## Summary
###

This tutorial followed the **tidycensus** chapters of [Analyzing US Census Data](https://walker-data.com/census-r/) by Kyle Walker. You built a Texas income map, a state education ranking, and a California age-population scatterplot.

### Exercise 1

Render and confirm all three figures appear. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 25)
```

###

<pre><code>---
title: "Analyzing US Census Data"
author: "&lt;your name&gt;"
execute:
  echo: false
---

&#96;&#96;&#96;{r}
#| message: false
library(tidyverse)
library(tidycensus)
library(sf)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| cache: true
income_tx <- get_acs(
  geography = "county",
  state = "TX",
  variables = "B19013_001",
  year = 2020,
  geometry = TRUE
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate), color = "white", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Median income") +
  labs(
    title = "Median household income by county, Texas (2020)",
    subtitle = "Suburban counties around major metros stand out from much of rural Texas",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
#| cache: true
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
edu_state |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .by = c(GEOID, NAME)
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent, fill = percent)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Adults with a bachelor's degree or higher by state-level geography (2020)",
    subtitle = "DC and several Northeastern states sit above most of the country",
    x = "State-level geography",
    y = "Share of adults 25 and over",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
age_ca <- get_acs(
  geography = "county",
  state = "CA",
  variables = c(median_age = "B01002_001", population = "B01003_001"),
  year = 2020,
  geometry = FALSE
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  mutate(label = if_else(min_rank(desc(population)) <= 8, NAME, "")) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point(color = "#2c7fb8", alpha = 0.75, size = 2.5) +
  geom_text(aes(label = label), hjust = -0.05, check_overlap = TRUE, size = 3) +
  scale_x_log10(
    labels = scales::label_number(scale_cut = scales::cut_short_scale()),
    expand = expansion(mult = c(0.02, 0.25))
  ) +
  labs(
    title = "Smaller California counties tend to be older (2020)",
    subtitle = "The largest counties are mostly younger than California's smallest counties",
    x = "Population (log scale)",
    y = "Median age",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
&#96;&#96;&#96;
</code></pre>

###

**tidycensus** accepts state names, postal codes, and FIPS codes. Changing the geography changes the units, not the basic workflow.

### Exercise 2

Publish your work to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If a render is still running, stop it first with `Ctrl/Cmd + C`. Copy and paste the resulting URL below.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 1)
```

###

Publishing turns your local analysis into a public page. That is how many real Census analyses reach readers.

### Exercise 3

Commit and push any remaining changes. Copy and paste the URL of your GitHub repo below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The workflow here is the modern Census stack: **tidycensus** for ACS data, **sf** for geometry, and **ggplot2** for display. Walker's book builds time-series comparisons, spatial statistics, and modeling on the same foundation.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
